One-arm spirals in locally isothermal, radially structured self-gravitating discs

### 2014-11-12 paper/results.tex
minor updates

Fig. \ref{2d_fargo_viz} shows the $m=1$ surface density
at the end of this simulation. For this one-armed spiral
we measure a co-rotation radius and growth rate  
\begin{align*}
  &R_c \simeq 4.4R_0,\\
  &\gamma\simeq 0.014\Omega_k(R_0) = 0.13\Omega_p. 
\end{align*}
The spiral can be considered as low frequency because its 
pattern speed $\Omega_p\lesssim 0.3\Omega$ in $R\in[R_{1},R_{2}]$,
where it is  most prominent. The growth rate $\gamma$ is also slow
relative to the local rotation. (Although the characteristic growth
time $\gamma^{-1} \simeq 10P_0$.) Thus, the spiral 
pattern is nearly stationary.

@@

Using the estimated value of $R_c$, we plot in
Fig. \ref{fargo_qbarrier} the quantity $\nu^2 - 1 + Q^{-2}$, which is
required to be positive in local theory for purely wave-like
solutions to the dispersion relation (Eq. \ref{dispersion}) when the
mode frequency is given. %Note that, with the above measured wavenumber $k$, 
%Eq. \ref{dispersion} requires $\nu^2 - 1 + Q^{-2}\sim0$ and the lower
%sign (long waves) to be taken. 
Fig. \ref{fargo_qbarrier} shows two $Q$-barriers located in
the inner disc, at $R_{Qb}=R_0$ and $R_{Qb}=1.6R_0$, which is indeed the
region where the $m=1$ spiral develops. This suggests that the one-arm
spiral is trapped. Note in this region, $\nu^2 - 1 +
Q^{-2}\simeq 0.1\ll 1$, as required for consistency with
Eq. \ref{dispersion} with the lower sign taken (i.e. long waves).  
There is one outer Lindblad resonance at $R_L\simeq
7.2R_0$. Thus, acoustic waves may be launched in $R\gtrsim 7.2R_0$ by
the spiral disturbance in the inner disc \citep{lin11b}. 
%implications: effect on dynamics in outer disc

>>>>

Fig. \ref{2d_fargo_viz} shows the $m=1$ surface density
at the end of this simulation. For this one-armed spiral
we measure a co-rotation radius and growth rate  
\begin{align*}
  &R_c \simeq 4.4R_0,\\
  &\gamma\simeq 0.014\Omega_k(R_0) = 0.13\Omega_p. 
\end{align*}
The spiral can be considered as low frequency because its 
pattern speed $\Omega_p \simeq 0.1\Omega_k(R_0)\lesssim 0.3\Omega$ in
$R\in[R_{1},R_{2}]$ (where it has the largest amplitude). The growth rate $\gamma$ is also slow
relative to the local rotation. (Although the characteristic growth
time $\gamma^{-1} \simeq 10P_0$.) Thus, the spiral 
pattern is nearly stationary.

@@

Using the estimated value of $R_c$, we plot in
Fig. \ref{fargo_qbarrier} the quantity $\nu^2 - 1 + Q^{-2}$, which is
required to be positive in local theory for purely wave-like
solutions to the dispersion relation (Eq. \ref{dispersion}) when the
mode frequency is given. %Note that, with the above measured wavenumber $k$, 
%Eq. \ref{dispersion} requires $\nu^2 - 1 + Q^{-2}\sim0$ and the lower
%sign (long waves) to be taken. 
Fig. \ref{fargo_qbarrier} shows two $Q$-barriers located in
the inner disc, at $R_{Qb}=R_0$ and $R_{Qb}=1.6R_0$; the bounded region is indeed the
where the $m=1$ spiral develops. This suggests that the one-arm 
spiral is trapped. Note in this region, $\nu^2 - 1 + 
Q^{-2}\simeq 0.1\ll 1$, as required for consistency with
Eq. \ref{dispersion} with the lower sign taken (i.e. long waves).  
There is one outer Lindblad resonance at $R_L\simeq
7.2R_0$. Thus, acoustic waves may be launched in $R\gtrsim 7.2R_0$ by
the spiral disturbance in the inner disc \citep{lin11b}. 
%implications: effect on dynamics in outer disc

### 2014-11-12 paper/summary.tex
minor updates

%non-sg sims, Be stars, eccentric discs
We stress that destabilisation through the temperature gradient
does not explicitly depend on self-gravity. In principle, this effect
will destabilise low-frequency $m=1$ trailing spirals regardless of
its origin. In our models, self-gravity coupled with the disc
structure permits a neutral one-arm mode, which is then destabilised. 
It would be of interest to explore other origins of the neutral
one-arm spiral. One possibility is in discs around classic Be stars
\citep{rivinius13}, for which several authors have demonstrated the
existence of one-arm oscillations \citep[see e.g.][and references
therein]{okasaki97,papaloizou06c,ogilvie08}. These authors employ
strictly isothermal disc models, so it would be interesting to explore
the effect of a temperature gradient on the stability of these
oscillations.

>>>>

%non-sg sims, Be stars, eccentric discs
We stress that destabilisation through the temperature gradient
does not explicitly depend on self-gravity. In principle, this effect
will destabilise low-frequency $m=1$ trailing spirals regardless of
its origin. In our models, self-gravity coupled with the disc
structure permits a neutral one-arm mode, which is then destabilised. 
It would be of interest to explore other origins of the neutral
one-arm spiral. One possibility is in Be star discs
\citep{rivinius13}, for which 
one-arm oscillations may explain long-timescale variations in their
emission lines \citep[see e.g.][and references
therein]{okasaki97,papaloizou06c,ogilvie08}. These studies employ
strictly isothermal disc models, so it would be interesting to explore
the effect of a temperature gradient on the stability of these
oscillations.
